### 2020-01-12 config/acronyms.tex

\usepackage[toc,acronym,section=section,nonumberlist,xindy]{glossaries}
\renewcommand*{\acronymname}{List of Abbreviations}

\renewcommand{\newacronym}[4][]{\newglossaryentry{#2}{type=\acronymtype,
name={#3},
short={#3},
long={#4},
shortplural={#3\glspluralsuffix},
longplural={#4\glspluralsuffix},
description={#4},
text={#3},
first={#4~(#3)},
plural={#3\glspluralsuffix},
firstplural={#4\glspluralsuffix~(#3\glspluralsuffix)},
#1}}

\newacronym{BLA}{BLA}{Best Linear Approximation}
\newacronym{MIMO}{MIMO}{Multiple-Input Multiple-Output}
\newacronym{SISO}{SISO}{Single-Input Single-Output}
\newacronym{PISPO}{PISPO}{Periodic-In Same Period-Out}
\newacronym{FRF}{FRF}{Frequency Response Function}
\newacronym{ADS}{ADS}{Advanced Design System}
\newacronym{pdf}{pdf}{probability density function}
\newacronym{PSD}{PSD}{Power Spectral Density}
\newacronym{RMS}{RMS}{Root Mean Square}
\newacronym{LTI}{LTI}{Linear Time-Invariant}
\newacronym{RF}{RF}{Radio Frequency}
\newacronym{WLAN}{WLAN}{Wireless Local Area Network}
\newacronym{EM}{EM}{electromagnetic}
\newacronym{DGS}{DGS}{Defected Ground Structure}
\newacronym{RMSE}{RMSE}{Root Mean Square Error}
\newacronym{TL}{TL}{transmission line}
\newacronym{MC}{MC}{Monte Carlo}
\newacronym{SC}{SC}{Stochastic Collocation}
\newacronym{VF}{VF}{Vector Fitting}
\newacronym{LHS}{LHS}{Latin Hypercube Sampling}
\newacronym{QMC}{QMC}{Quasi Monte Carlo}
\newacronym{RBF}{RBF}{Radial Basis Functions}
\newacronym{PCBs}{PCBs}{Printed Circuit Boards}
\newacronym{PDNs}{PDNs}{Power Distribution Networks}
\newacronym{LDS}{LDS}{Low Discrepancy Sampling}
\newacronym{LBTLS}{LBTLS}{Local Bootstrapped Total Least Squares}
\newacronym{VNA}{VNA}{Vector Network Analyzer}
\newacronym{FOMs}{FOMs}{Figures of Merit}
\newacronym{FOM}{FOM}{Figure of Merit}
\newacronym{SNR}{SNR}{Signal-to-Noise Ratio}
\newacronym{SINAD}{SINAD}{Signal-to-Noise and Distortion ratio}
\newacronym{ACPR}{ACPR}{Adjacent Channel Power Ratio}
\newacronym{EVM}{EVM}{Error Vector Magnitude}
\newacronym{NPR}{NPR}{Noise Power Ratio}
\newacronym{DUT}{DUT}{Device Under Test}
\newacronym{TOI}{TOI}{Third Order Intercept point}
\newacronym{OFDM}{OFDM}{Orthogonal Frequency Division Multiplexing}
\newacronym{LTE}{LTE}{Long Term Evolution}

% \makeglossaries
\makenoidxglossaries