### 2020-03-20 Gus-thesis.tex

%%
%% This code was generated from the thesis template of Adam Cooman--Evi Van Nechel.
%% It has been adapted slightly by Gustavo Quintana.
%% Compile with: 
%% cd C:\GusDocs\org\thesis\Gus-thesis
%% pdflatex --shell-escape -synctex=1 -interaction=nonstopmode Gus-thesis.tex 
%%
\documentclass[english, british, BCOR=5mm, DIV=12, fontsize=10pt]{scrbook}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[paperwidth=17cm, paperheight=24cm]{geometry} % \usepackage[hmargin=1.5cm, textheight=25cm]{geometry}
\geometry{verbose, tmargin=2.5cm, bmargin=3cm, lmargin=2.5cm, rmargin=2cm}
\setcounter{secnumdepth}{2}
\setcounter{tocdepth}{2}
\setlength{\parskip}{\medskipamount}
\setlength{\parindent}{0pt}
\synctex=-1
\usepackage{color}
\definecolor{note_fontcolor}{rgb}{0.800781, 0.800781, 0.800781}
\usepackage{amsfonts, amsmath, amssymb, amsthm, amsbsy, array, babel, booktabs, bm, float, graphicx, lineno, mathdots, mathptmx, mathtools, newproof, pgf, subcaption, textcomp, times, units, verbatim, xcolor}
\usepackage[unicode=true, pdfusetitle, bookmarks=true, bookmarksnumbered=false, bookmarksopen=false, breaklinks=true, pdfborder={0 0 0}, pdfborderstyle={}, backref=false, colorlinks=false]{hyperref}
\usepackage[outdir=./]{epstopdf}
\usepackage{tikz, circuitikz}
\usetikzlibrary{shapes, arrows, calc, patterns, decorations.pathmorphing, decorations.markings, positioning}

\newtheorem{thm}{Theorem}
\newtheorem{lem}[thm]{Lemma}
\newproof{pf}{Proof}

\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}

\usepackage{algorithm, algorithmic}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}
%% The greyedout annotation environment
\newenvironment{lyxgreyedout}
  {\textcolor{note_fontcolor}\bgroup\ignorespaces}
  {\ignorespacesafterend\egroup}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\input{config/preamble.tex}

% to stop the compiler from complaining about the \it command in the references, I override it with the thing KOMA proposes.
\renewcommand{\it}{\normalfont\itshape}

% lyxgreyedout defined in latex
\renewenvironment{lyxgreyedout}
{\textcolor{red}\bgroup}{\egroup}

%\hyphenation{Lon-g-las-t-na-me}
\usepackage[square]{natbib}
\bibliographystyle{plainnat} % wmaainf  apalike

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Start Document
\makeatother

\begin{document}
{\footnotesize{}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  Front cover
%\includepdf[pages={1}]{Cover}
\newpage{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% VUB title page

\titlepage{\fontfamily{pag}\selectfont % I use the ugly VUB font only for this title page
	\hspace{-0.1cm}\includegraphics[width=0.4\columnwidth]{Extras/VUB_logo.pdf}
	% \vspace{-1.7cm}
	\hfill{}%
	\begin{minipage}[t]{0.4\columnwidth}% 
	\vspace{-1.4cm}
		\begin{flushright}
			{\footnotesize{}Department of }\\
			{\footnotesize{}Fundamental Electricity}\\
			{\footnotesize{}and Instrumentation (ELEC)}
		\par\end{flushright}%
	\end{minipage}

	\vspace{1.5cm}
	% I use purple and orange here for the color. In the preamble I have overwritten these colors to be VUB orange and VUB blue respectively
	\textcolor{purple}{\footnotesize{}Thesis submitted in fulfillment of the requirements for the degree of}\\
	\textcolor{purple}{\footnotesize{}Doctor in Engineering (Doctor in de Ingenieurswetenschappen) by}{\footnotesize\par}
	\textcolor{orange}{\Large{}ir. Gustavo Quintana Carapia}{\Large\par}

	\vspace{1cm}

	\begin{flushleft}
		\textcolor{purple}{\LARGE\noindent%\textcolor{purple}
		\textbf{Statistical analysis and experimental validation of data-driven dynamic measurement methods}}\\
	\par\end{flushleft}

	\vfill{}

	\begin{tabular}{@{}lll@{}}
	\textbf{\small PROMOTERS}		& &	{\small prof. dr. ir. Ivan Markovsky}				\\
				& &		{\footnotesize{}ELEC Department, Vrije Universiteit Brussel, Belgium}\\[0.3em]
				& &	\small prof. dr. ir. Rik Pintelon				\\
				& &		{\footnotesize{}ELEC Department, Vrije Universiteit Brussel, Belgium}	\\[0.3em]
				\\
	\textbf{\small MEMBERS OF THE JURY}	&	&\small prof. dr. ir. Patrick Guillaume (President)		\\
				&	&{\footnotesize{}MECH Department, Vrije Universiteit Brussel, Belgium}	\\[0.3em]
				% &								\\
				&	&\small prof. dr. ir. Roger Vounckx (Vice-President)				\\
				&	&{\footnotesize{}ETRO Department, Vrije Universiteit Brussel, Belgium}	\\[0.3em]
				% &								\\
				&	&\small dr. ir. Philippe Dreesen (Secretary)					\\
				&	&{\footnotesize{}ELEC Department, Vrije Universiteit Brussel, Belgium}				\\[0.3em]
				% &								\\
				&	&\small dr. ir. Nikolaos Deligiannis \\
				&	&{\footnotesize{}ETRO Department, Vrije Universiteit Brussel, Belgium} 			\\[0.3em]
				% &								\\
				&	&\small prof. dr. ir. Lyudmila Mihaylova 		\\
				&	&{\footnotesize{}ACSE Department, University of Sheffield, United Kingdom} \\
								% &								\\
				&	&\small prof. dr. ir. Stephane Chretien  		\\
				&	&{\footnotesize{}ERIC Laboratory, University Lyon 2, France} \\
								% &								\\
				&	&\small prof. dr. ir. Guillaume  Merc\`ere 		\\
				&	&{\footnotesize{}LIAS Laboratory, Universit\'e de Poitiers, France} \\
\end{tabular}

}

\newpage{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  Book details

\pagenumbering{gobble}
~\vfill{}

% {\footnotesize{}ISBN: } \\% You need to apply online for an ISBN, this takes a couple of days
{\footnotesize{}DOI: 10.13140/RG.2.2.27482.29129}\\ % I uploaded my thesis to researchgate to generate a DOI: if you remove it from ResearchGate, the DOI will no longer exist!
{\footnotesize{}~}{\footnotesize\par}

{\footnotesize{}University Press BVBA}\\
{\footnotesize{}Rechtstro 2/001, 9185 Wachtebeke, Belgium }\\
{\footnotesize{}https://www.universitypress.be/}\\
{\footnotesize{}~}{\footnotesize\par}

{\footnotesize{}Vrije Universiteit Brussel, dept. ELEC}\\
{\footnotesize{}Pleinlaan 2, 1050 Brussel, Belgium}\\
{\footnotesize{}http://vubirelec.be/}\\
{\footnotesize{}~}{\footnotesize\par}

%{\footnotesize{}Cover design by ...}\\
%{\footnotesize{}Background image by ...}{\footnotesize\par}
%{\footnotesize{}~}{\footnotesize\par}

{\footnotesize{}$\copyright$ February 2020, Gustavo Quintana Carapia.}\\
{\footnotesize{}~}{\footnotesize\par}

{\footnotesize{}All rights reserved. No parts of this document may be reproduced or transmitted in any form or by any means without the prior written permission of the authors.}{\footnotesize\par}

{\footnotesize\par}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  Acknowledgements
\newpage{}
% \shipout\null
\pagenumbering{roman}

%\include{Extras/Acknowledgements}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  TOC and abbreviations list
\tableofcontents{}

% table of glossaries didn't work so I made a list of abbreviations manually
\include{Extras/Abbreviations}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  Thesis
% CustomCommands.lyx contains my custom lyx math macros.
\include{config/CustomCommands}

\include{Extras/Summary}

\pagenumbering{arabic}

\include{ChapterIntroduction/ChapterIntroduction}

%\part{Model driven filter design}\label{part:Filter}

\include{ChapterPreliminaries/ChapterPreliminaries}

\include{ChapterStatisticalAnalysis/ChapterStatisticalAnalysis}

\include{ChapterExperimentalValidation/ChapterExperimentalValidation}

\include{ChapterRampInput/ChapterRampInput}

%\part{Figures of merit for~characterizing nonlinear~devices}\label{part:PA}

\addtocontents{toc}{\vspace{2em}} \include{ChapterConclusion/ChapterConclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  Start appendix

\appendix
\addtocontents{toc}{\vspace{0em}} \include{ChapterAppendices/ChapterAppendices}

\include{Extras/ListofPublications}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  Bibliography

\vspace{-0.5cm}

\addcontentsline{toc}{chapter}{Bibliography}

\bibliography{Gus-thesis}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  Back cover
\newpage{}
\shipout\null
%\includepdf[pages={2}]{Cover}

\end{document}