### 2020-08-21 Extras/ListofPublications.tex

%!TEX root = ..\Thesis.tex

\addchap{List of Publications}

\subsubsection*{Journal publications}

\noindent G. Quintana-Carapia, I. Markovsky, R. Pintelon, P. Z. Csurcsia and D. Verbeke, "Bias and covariance of the least squares estimate in a structured errors-in-variables problem," \textit{Computational Statistics & Data Analysis}, Vol. 144, No. 106893, 2020, ISSN 0167-9473, doi: 10.1016/j.csda.2019.106893.

%\vspace{0.1618cm}

\noindent G. Quintana-Carapia, I. Markovsky, R. Pintelon, P. Z. Csurcsia and D. Verbeke, "Experimental validation of a data-driven step input estimation method for dynamic measurements," \textit{IEEE Transactions on Instrumentation and Measurement}, \color{blue} vol. 69, no. 7, pp. 4843-4851, July 2020, \color{black} doi: 10.1109/TIM.2019.2951865.

%\vspace{0.1618cm}

\noindent G. Quintana-Carapia, I. Markovsky, "Input parameters estimation from time-varying measurements," \textit{Measurement}, Vol. 153, No. 1, pp. 107418, 2020, ISSN 0263-2241, doi: 10.1016/j.measurement.2019.107418.

\vspace{0.1618cm}
\subsubsection*{Conference publications}

\noindent G. Quintana-Carapia, I. Markovsky, "Data driven dynamic measurements",
In: \textit{9th International Workshop on the Analysis of Dynamic Measurements}", Berlin, Germany, 2016.

%\vspace{0.1618cm}

\noindent G. Quintana-Carapia, I. Markovsky, "Data driven dynamic measurements",
In: \textit{35th Benelux Meeting on Systems and Control}", Soesterberg, The Netherlands, 2016.

%\newpage